Reproducibility and Open Science in Economics

### 2025-02-09 article.tex
Updates from Overleaf

Many well-known surveys impose redistribution and usage restrictions that are not consistent with the OKD definition. Many such redistribution restrictions apply to datasets with more detailed personal information collected through surveys, and are meant to ensure continued compliance with ethical rules of behavior, often with informed consent agreements by survey participants and local privacy laws. Notable examples include  PSID \parencite{institute_for_social_research_panel_2024}, World Value Survey \parencite{haerpfer_world_2024}, Demographic Household Surveys \parencite{dhs_program_demographic_2024}, German Socio-Economic Panel \parencite{goebel_german_2019,goebel_socio-economic_2024}, all of which have broad (cost-free) usage, conditional on registration and compliance with usage restrictions. Table~\ref{}

@@

@@

The core of a paper's analysis, however, resides in how the more general software package is manipulated, in combination with the data, via \textbf{instructions}. I do not call these `source code', as that terms tends to be used in conjunction with complex compiled software, such as those listed as \textbf{high-level software}. Rather, these instructions are interpreted code, or compiled just-in-time, in the language used by the high-level software, or, as is still sometimes the case, in the form of instructions to humans on how to manipulate a non-programmed user interface for software.\footnote{In economics, this appears to occur most often for \ac{GIS} software, and sometimes for data extraction software used by data providers.}

@@

@@

In the template README published by myself and several other data editors in economics \parencite{templateREADMEv1.1}, we emphasize that a README should provide enough instructions for a reasonable person to re-implement the analysis described in the replication package. Here, authors need to take into account that cutting edge methods, including technology, may require more information and instruction than for standard methods. Authors likely do not need to describe how to run an analysis in Stata, given its ubiquitous use in economics, and the ease with which instructions can be found more generally, but they may need to provide detailed step-by-step information if the technology used is rare or bespoke. For instance, the emerging use of \acp{LLM} and \ac{AI} methods is far from the economic mainstream as of the writing of this article. Recent articles are still identifying ways economists scan actually use these tools, both for personal productivity \parencite{korinek_generative_2023} and as part of the technical toolkit \parencite{athey_machine_2019,dell_deep_2024}. Only one of the articles submitted to any of the AEA journals in 2024 used these techniques (to be named, not yet published).

@@

Some newer research requires vastly different types of resources. Studies using raw satellite data may require more than 10TB of data storage \citep{khachiyan_using_2022,khachiyan_data_2022}, may need more than 20,000 compute hours on a cloud provider \citep{rudik_optimal_2020,rudik_data_2020}, or the use of one \citep{dell_deep_2024,dell_data_2025} or dozens \citep{khachiyan_data_2022} \acp{GPU}.\footnote{As of January 2025, the type of \ac{GPU} used by \citet{dell_data_2025}, costs between USD 4500 and USD 7700, or between 2 and 7 times as much as a standard laptop.} But  access to the code and data for these papers is open: The AEA-related replication packages for these articles are licensed under a standard \ac{CC-BY} license. Some of the data not included in the \citet{dell_data_2025} replication package is on Huggingface, also under a \ac{CC-BY} license \citep{dell_research_harvard_newswire_nodate}.

Are such computational constraints a problem? No consistent analysis exists that correlates resource requirements to academic outcomes such as citations, primarily because it is very hard to measure consistently the resource requirements of economic articles. The very small sample in the previous paragraph may serve to illustrate this, but without controls for scientific merit, is purely an indicator. \citet{dell_research_harvard_newswire_nodate} had been downloaded 98 times in December 2024, six months after the arXiv paper associated with it was published \citep{silcock_newswire_2024}. \citet{rudik_data_2020} has had  1432 views, 124 downloads for replication package, as the manuscript \citep{rudik_optimal_2020} has 15 citations. The replication package \citet{khachiyan_data_2022} has  2124 views, 175 downloads, while the manuscript \citep{khachiyan_using_2022} has 5 citations (all as of January 2025). For comparison, the average article in one of the \ac{AEA}'s journals has 908 views and 106 downloads \citep[Table 4][]{vilhuber_report_2025}.

@@

>>>>

Many well-known surveys impose redistribution and usage restrictions that are not consistent with the OKD definition. Many such redistribution restrictions apply to datasets with more detailed personal information collected through surveys, and are meant to ensure continued compliance with ethical rules of behavior, often with informed consent agreements by survey participants and local privacy laws. Notable examples include  PSID \parencite{institute_for_social_research_panel_2024}, World Value Survey \parencite{haerpfer_world_2024}, Demographic Household Surveys \parencite{dhs_program_demographic_2024}, German Socio-Economic Panel \parencite{goebel_german_2019,goebel_socio-economic_2024}, all of which have broad (cost-free) usage, conditional on registration and compliance with usage restrictions. Table~\ref{

@@

% table of data access in AER in 2024 using internal categorization

@@

The core of a paper's analysis, however, resides in how the more general software package is manipulated, in combination with the data, via \textbf{instructions}. I do not call these `source code', as that terms tends to be used in conjunction with complex compiled software, such as those listed as \textbf{high-level software}. Rather, these instructions are mostly interpreted code, or compiled just-in-time, in the language used by the high-level software, or, as is still sometimes the case, in the form of instructions to humans on how to manipulate a non-programmed user interface for software.\footnote{In economics, this appears to occur most often for \ac{GIS} software, and sometimes for data extraction software used by data providers.}

@@

The second channel is through the \textbf{instructions}. Conditional on having the right software and the appropriate resources (see next section), most economics papers have openly accessible code. The advent of data and code availability policies most certainly has helped, but also reflected attitudes already present in the researcher community.\footnote{For a review of the history of data and code availability policies in economics, see \citet{vlaeminck_dawning_2021}.} While not the first, the AEA's policy was first implemented in 2005 \citep{bernanke_editorial_2004,american_economic_association_data_2005}, but the oldest replication package curated by the AEA accompanied an 1999 article \citep{frankel_does_1999,frankel_replication_1999} preceding the policy by several years.\footnote{I believe the oldest replication package in economics may be \citet{koenker_asymptotic_1988-1}, see \url{https://aeadataeditor.github.io/posts/2023-02-02-oldest-replication-package-jae}.} While the focus of availability policies was on the data, the code often came along for the ride, albeit not always in its most complete form. The advent of newer data and code availability policies \citep[for the AEA, ][]{AEA-announcement-July-2019,AEA-announcement-July-2019}, culminating for now in multiple journals having policies compliant with \ac{DCAS} \citep{koren_data_2022}, such as the AEA \citep{american_economic_association_data_2024} and

@@

In the template README published by myself and several other data editors in economics \parencite{templateREADMEv1.1}, we emphasize that a README should provide enough instructions for a reasonable person to re-implement the analysis described in the replication package. Here, authors need to take into account that cutting edge methods, including technology, may require more information and instruction than for standard methods. Authors likely do not need to describe how to run an analysis in Stata, given its ubiquitous use in economics, and the ease with which instructions can be found more generally, but they may need to provide detailed step-by-step information if the technology used is rare or bespoke. For instance, the emerging use of \acp{LLM} and \ac{AI} methods is far from the economic mainstream as of the writing of this article. Recent articles are still identifying ways economists scan actually use these tools, both for personal productivity \parencite{korinek_generative_2023} and as part of the technical toolkit \parencite{athey_machine_2019,dell_deep_2024}. 
%Only one of the articles submitted to any of the AEA journals in 2024 used these techniques (to be named, not yet published).

@@

Some newer research requires vastly different types of resources. Studies using raw satellite data may require more than 10TB of data storage \citep{khachiyan_using_2022,khachiyan_data_2022}, may need more than 20,000 compute hours on a cloud provider \citep{rudik_optimal_2020,rudik_data_2020}, or the use of one \citep{dell_deep_2024,dell_data_2025} or dozens \citep{khachiyan_data_2022} \acp{GPU}.\footnote{As of January 2025, the type of \ac{GPU} used by \citet{dell_data_2025}, costs between USD 4500 and USD 7700, or between 2 and 7 times as much as a standard laptop.} But  access to the code and data for these papers is open: The AEA-related replication packages for these articles are licensed under a standard \ac{CC-BY} license. Some of the data not included in the \citet{dell_data_2025} replication package is on Huggingface, also under a \ac{CC-BY} license \citep{silcock_newswire_2024}.

Are such computational constraints a problem? No consistent analysis exists that correlates resource requirements to academic outcomes such as citations, primarily because it is very hard to measure consistently the resource requirements of economic articles. The very small sample in the previous paragraph may serve to illustrate this, but without controls for scientific merit, is purely an indicator. \citet{silcock_newswire_2024} had been downloaded 98 times in December 2024, six months after the arXiv paper associated with it was published \citep{silcock_newswire_2024}. \citet{rudik_data_2020} has had  1432 views, 124 downloads for replication package, as the manuscript \citep{rudik_optimal_2020} has 15 citations. The replication package \citet{khachiyan_data_2022} has  2124 views, 175 downloads, while the manuscript \citep{khachiyan_using_2022} has 5 citations (all as of January 2025). For comparison, the average article in one of the \ac{AEA}'s journals has 908 views and 106 downloads \citep[Table 4]{vilhuber_report_2025}.

@@

\section{The Benefits of Open Science in Economics}
\label{sec:benefits}

To illustrate some of the benefits of open science, I describe two recent articles in economics that leverage the availability of code and data to improve econometric methods for future researchers. The first paper relies on the empirical recomputation of prior papers to assess a theoretically ambiguous potential bias in inference. The second paper also focuses on inference, and uses actual data from previous papers to simulate the relevance of the impact. Both then provide new software (R and Stata packages), under open source licenses, to ``fix'' the problem for future studies.

\citet{roth_pretest_2022}  uses 12 previously published papers to assess whether the usual tests for pre-existing differences in trends when using difference-in-differences methods are properly controlling for power, and the empirical impact on subsequent inference. The theoretical bias is ambiguous, so an empirical evaluation is necessary. The study both leverages the open availability of materials in economic journals, but also illustrates the limitations imposed by imperfect adherence to openness. To wit, Roth writes that he searched for

\begin{quote}
``the phrase “event study” in papers published in the \textit{American Economic Review, American Economic Journal: Applied Economics}, and \textit{American Economic Journal: Economic Policy} between
2014 and June 2018 ... The search returned 70 total papers that include a figure that the authors describe as an event-study plot.''
\end{quote}

but continues to then be limited by lack of data in the majority of cases:

\begin{quote}
    ``I exclude 43 papers for which data to replicate the main event-study plot were unavailable. \citep[pg. 307]{roth_pretest_2022}''\footnote{He also excludes another 15 papers for reasons not related to data availability.}
\end{quote}

While it remains unclear whether the excluded papers are non-compliant with the AEA's policy at the time, or whether they have legitimate reasons not to provide the data (Roth does not provide the raw result of his search), the paper is able to make an important methodological point (723 citations as of January 2025, per Google Scholar) because it is able to fully recompute the results in previous papers, apply new tests and methodologies, and come to meaningful recommendations and tools --- Roth provides an (open source) R package to implement his methodology.

\citet{de_chaisemartin_at_2024} use open access information on RCTs (AEA registry) to find 15 RCTs of a particular type (clustered paired or small strata), of which 4 have publicly available data (and reproducible artifacts). They then provide results both on simulations using these data, and in particular, re-estimate the regressions used in those studies and apply their proposed solution, showing that the number of significant effects is reduced by one-third. In other work \citep{de_chaisemartin_two-way_2020,de_chaisemartin_difference--differences_2024}, the results from various other papers are also recomputed to empirically demonstrate the relevance of the proposed methods, and software packages \citep[e.g.][]{de_chaisemartin_chaisemartinpackagesdid_multiplegt_dyn_2025} are developed and made openly available.\footnote{Note that as of January 2025, many of the packages do not have an explicit open source license --- or any license --- applied, a common feature of economists working in the open source world. My presumption is that they simply assume that everybody knows that the code is openly available.}

In both of these examples, the ability to access prior data, code, and information is critical to improving future scientific progress, but is limited by both historical and unavoidable limitations on openness.
